
# __Chapter 3: Logistic Regression__

## __3.1 Create a new R Markdown file__ 

```{r}
# File 'chapter3.Rmd' is created and linked to index 
# access the tidyverse libraries tidyr, dplyr, ggplot2 
library(readr); library(dplyr); library(ggplot2); library(tidyr); library(gtsummary)
```

## __3.2 Read Data Files__
```{r}

# Read the CSV file 
alc<- read_csv("data/alc.csv")
colnames(alc)
```
### About the data
This data is obtained from two secondary education schools in Portugal. it was collected by using school reports and questionnaires.

##### __The data include the following__ : 
1. __School name__ 
2. __Demographic data__ (Sex, age, address)
3. __Socio-economic Status__( family size, Parent status, Parents education and occupation, and guardians)
4. __Social Data__ (extra-curricular activities, romantic relationship, family relationship, free time and going out with friends)
5. __School related feature__ :(travel time to school, study time/week , extra support from school and family, extra-paid classe and abscences)
6. __Health status__
7. __Alcohol consumption__ : Alcohol consumption in weekdays and weekends. With Average of consumption/week, and categorize them as high consumption (>2 times)
6. __Performance__ : It include past class failures and the students' performance in Mathematics and Portuguese language in 3 period  G1 and G2 correspond to the 1st and 2nd period grades and G3 for the final year grade.



## __3.3 Relation between Alcohol consumption and 4 Variables__ 

```{r}
# Create a logistic regression model
reg_model <- glm(high_use ~  sex +failures + absences + G3, data = alc, family = "binomial")

# print summary of the model
summary(reg_model)

# Odds Ratio(OR) computation from coefficients
OR<-coef(reg_model) %>% exp

# compute confidence intervals (CI)
CI<- confint(reg_model)%>% exp

# print out the odds ratios with their confidence intervals
cbind(OR, CI)


```
The variables selected 
1. Gender: Hypothesis from student's sex we can predict the high alcohol consumption.
(hypothesis that males will be more alcohol consumer than females)
2. Health: Hypothesis from the students' health we can predict the high alcohol consumption. (hypothesis that high-use group will have lower health score than low-use group)
3. Absences: from the students' grades we can predict the high alcohol consumption
(hypothesis that high-use group will have more absences than low-use group)
4. G3: Hypothesis from the students' grades we can predict the high alcohol consumption . (hypothesis that high-use group will have lower score than low-use group)

## __3.4 Numerical and Graphically explore variables__ 
#### __3.4.1. Gender relationships with alcohol consumption__
```{r}
#Bars of high_use and Sex
ggplot(alc, aes(x = high_use,  fill = sex))+ geom_bar() + xlab("Gender")+  ggtitle("Relation between Alcohol Consumption and Gender")+ theme(plot.title = element_text(hjust = 0.5 , size=15, face="bold", color = "blue"))

#Summarize high alcohol consumption by gender
 alc %>% select(high_use,sex) %>% tbl_summary(by=c("high_use"),percent = "row")

```

* According to gender : the males with high alcohol consumption (N=70) are more than females (N=42). 
* The results revealed that 40% of males are high alcohol consumption while only 21% of females are high alcohol consumption.
* I accept my hypothesis



#### __3.4.2. Health relationships with alcohol consumption__
```{r}

#  Boxplot of high_use and Health
 ggplot(alc, aes(x = high_use, y = health, col = high_use))+ geom_boxplot() +  ggtitle("Relation between Alcohol Consumption and health")+ theme(plot.title = element_text(hjust = 0.5 , size=15, face="bold", color = "blue"));

ggplot(alc, aes(x = high_use, y = health, col = sex))+ geom_boxplot() +  ggtitle("Relation between Alcohol Consumption and health in both genders")+ theme(plot.title = element_text(hjust = 0.5 , size=15, face="bold", color = "blue"))

#Summarize relation between high alcohol consumption and health in  both genders 
alc %>% group_by(sex, high_use) %>% summarise(mean(health),sd(health));
alc %>% group_by( high_use) %>% summarise(mean(health),sd(health))
```
* The overall health score for high alcohol (3.73) is higher than low consumption (3.49) * The results showed that the mean health score for high-use alcohol in males is more than females (the mean health score for females=3.39 , males= 3.92). 
* In Males the  the difference between low alcohol and high alcohol consumption students was higher in males(low=3.67, high= 3.93)
* In Females group: the difference between low alcohol and high alcohol consumption students was small (low=3.37, high=3.39)

* The results is different from my hypothesis


#### __3.4.3. Absences relationships with alcohol consumption__
```{r}


 #  Boxplot of high_use and Absences
 ggplot(alc, aes(x = high_use, y = absences, col = high_use))+ geom_boxplot() +  ggtitle("Relation between Alcohol Consumption and Abscences")+ theme(plot.title = element_text(hjust = 0.5 , size=15, face="bold", color = "blue"));
 ggplot(alc, aes(x = high_use, y = absences, col = sex))+ geom_boxplot() +  ggtitle("Relation between Alcohol Consumption and Abscences")+ theme(plot.title = element_text(hjust = 0.5 , size=15, face="bold", color = "blue"))

#Summarize relation between high alcohol consumption and absences in  both genders 
alc %>% group_by(sex, high_use) %>% summarise(mean(absences),sd(absences));
alc %>% group_by( high_use) %>% summarise(mean(absences),sd(absences))


```
* The overall abscences in high-use is higher than the low-use group (low=3.7, high=6.4)
* The mean students' abscences in school among the high-use 6.85 in females and 6.1 in males 
* I accept my hypothesis


#### __3.4.3. Grades relationships with alcohol consumption__
```{r}

#  Boxplot of high_use and G3
 ggplot(alc, aes(x = high_use, y = G3, col = high_use))+ geom_boxplot() + ylab("Grades")+  ggtitle("Relation between Alcohol Consumption and Grades")+ theme(plot.title = element_text(hjust = 0.5 , size=15, face="bold", color = "blue"));
ggplot(alc, aes(x = high_use, y = G3, col = sex))+ geom_boxplot() + ylab("Grades")+  ggtitle("Relation between Alcohol Consumption and Grades in both Genders")+ theme(plot.title = element_text(hjust = 0.5 , size=15, face="bold", color = "blue"))

#Summarize relation between high alcohol consumption and Grades in  both genders 
alc %>% group_by(sex, high_use) %>% summarise(mean(G3),sd(G3));alc %>% group_by( high_use) %>% summarise(mean(G3),sd(G3))
```
* The mean grades for low alcohol consumption is 11.8 while the high alcohol consumption is 10.9 . 
* The males showed a large difference between both consumption groups (low=12.3, high=10.3) While the females there was only 0.4 difference (low=11.4 , high=11.8)
* I accept my hypothsis


## __3.5 Create Logistic regression __ 

```{r}
# Create a logistic regression model
reg_model <- glm(high_use ~  sex +health + absences + G3, data = alc, family = "binomial")

# print summary of the model
summary(reg_model)

# Odds Ratio(OR) computation from coefficients
OR<-coef(reg_model) %>% exp

# compute confidence intervals (CI)
CI<- confint(reg_model)%>% exp

# print out the odds ratios with their confidence intervals
cbind(OR, CI)
```
* The regression model showed statistically significant variables in Gender, abscence and Grades. 
* From the model the following show higher probability of high alcohol consumption :

> 1. Gender is Males (OR= 2.74)
> 2. Better Health score (OR= 1.07) But it is not signficant
> 3. More absence (OR= 1.09)
> 4. Lower grade (OR= 0.93)

* the result are the same as previous stated hypothesis except for the health which is not significant predictor for consumption of alcohol among students


## __3.6 Prediction__ 

```{r}

# according to your logistic regression model, had a statistical relationship with high/low alcohol consumption, explore the predictive power of you model

# Predict probability of high_use and add it to alc
probabilities <- predict(reg_model, type = "response")
alc <- mutate(alc, probability = probabilities)

# use the probabilities to make a prediction of high_use
alc <- mutate(alc, prediction = probability > 0.5)

# Provide a 2x2 cross tabulation of predictions versus the actual values and optionally display a graphic visualizing both the actual values and the predictions

# tabulate the target variable versus the predictions
table(high_use = alc$high_use, prediction = alc$prediction)

#plot of 'high_use' versus 'probability' in 'alc'
ggplot(alc, aes(x = probability , y = high_use, col=prediction)) +geom_point()


#Compute the total proportion of inaccurately classified individuals (= the training error) and comment on all the results. Compare the performance of the model with performance achieved by some simple guessing strategy  #line 700 in excercise
# define a loss function (mean prediction error)
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}

# call loss_func to compute the average number of wrong predictions in the (training) data
loss_func(class = alc$high_use, prob = alc$probability)

```

* The average number of wrong prediction in the training data is 25.7% . So this is a good model for prediction of alcohol consumption among students. 

