
# Chapter 6: Analysis of Longitudinal data

## 6.1 Meet and Repeat: PART I but using the RATS data 
*  the analyses of Chapter 8 of MABS, using the R codes of Exercise Set 6: Meet and Repeat: PART I but using the RATS data (from Chapter 9 and Meet and Repeat: PART II). 
(0-7 points: 0-4 points for graphs or analysis results + 0-3 points for their interpretations)

```{r}
# Load packages 
library(readr)
library(dplyr)
library(ggplot2)

RATSL <- read.csv("data/RATSL.csv")


glimpse(RATSL)

# Convert categorical variables in RATS to factors.
RATSL$ID <- factor(RATSL$ID)
RATSL$Group <- factor(RATSL$Group)

# Glimpse the data
glimpse(RATSL)
```
 
 
### Graphical Presentation 
```{r}

# Draw RATSL plot
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:8, times=2)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)))

```


 
### Standardizing Data
```{r}
# standardise RATSL variable 
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate(stdWeight = (Weight - mean(Weight))/sd(Weight)) %>%
  ungroup()

# Glimpse the data
glimpse(RATSL)


# Plot standardized RATSL
 ggplot(RATSL, aes(x = Time, y = stdWeight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:8, times=2)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardized RATSL")

```
Interpretation 
* Standardization reduces weight variation. The follow-up of all treatment groups. Group 2 and Group 3 gain weight more than in group 1. Group 2 show outlier in its readings

### summary RATSL with mean and standard error of the variable

```{r}
n=8
RATSS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise(mean = mean(Weight), se = sd(Weight)/sqrt(n)) %>%
  ungroup()

# plot the mean profiles
ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")
 

```
Interpretation 
The standard error of mean is shown in this graph which showed  the heighest variance in group 2 and the lowest for group 1 . The group 3 SE is decreasing by time. 
### 
```{r}

# # Create a summary data with mean (ignoring baseline )
 RATSLX <- RATSL %>%
  filter(Time > 1) %>%
  group_by(Group, ID) %>%
  summarise(mean=mean(Weight)) %>%
  ungroup()

# summary of the RATSLX
summary(RATSLX)

# box plot for RATSLX
ggplot(RATSLX, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "(Weight)/ days ")
```

#### Remove outliers
```{r}

# Comment
 # create a new data by removing the oulier 
RATSLX2 <- RATSLX %>% filter(mean < 550)

# draw a boxplot of without the outliers 
ggplot(RATSLX2, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "(Weight)/ days ")
```


### analysis of variance (ANOVA)
```{r}

# Fit the linear model with the mean as the response 
fit <- lm(mean ~ Group, data = RATSLX2)
 anova(fit)

```
Interpretation 
ANOVA test showed that there is highly signficant difference in the weight of the 3 study groups


## 6.2 Meet and Repeat: PART II  using the BPRS data

```{r}
# Read BPRSL dataset
BPRSL <- read.csv("data/BPRSL.csv")

# Factor treatment & subject
BPRSL$treatment <- factor(BPRSL$treatment)
BPRSL$subject <- factor(BPRSL$subject)

str(BPRSL)
```
text continues here.

#### Standardize BPRS
```{r}

# Standardise the variable bprs
BPRSL <- BPRSL %>%
  group_by(week) %>%
  mutate(stdbprs = (bprs-mean(bprs))/sd(bprs)) %>%
  ungroup()
# Glimpse the data
glimpse(BPRSL)

# Create a summary data by treatment and subject with mean as the summary variable (ignoring baseline week 0)
BPRSL8S <- BPRSL %>%
  filter(week > 0) %>%
  group_by(treatment, subject) %>%
  summarise( mean=mean(bprs) ) %>%
  ungroup()

# Glimpse the data
glimpse(BPRSL8S)

# Draw a boxplot of the mean versus treatment
library(ggplot2)
ggplot(BPRSL8S, aes(x = treatment, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(bprs), weeks 1-8")



```
 
#### Create Regression model for BPRS
```{r}

# create a regression model for BPRS
BPRS_reg <- lm(bprs ~ week + treatment, data = BPRSL)


# access library lme4
library(lme4)

# Create a random intercept model
BPRS_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)


# summary of the intercept model 
summary(BPRS_ref)

```
text 
#### Title
```{r}

# create a random intercept and random slope model
BPRS_ref2 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)

#summary of model 
summary(BPRS_ref2)

```

#### ANOVA for both models
```{r}
# perform an ANOVA test on the two models
anova(BPRS_ref2, BPRS_ref)


 
```
ANOVA test showed a statistically signficance across time 
#### Create random intercept and random slope with interaction
```{r}
# create a random intercept and random slope model with the interaction
BPRS_ref3 <- lmer(bprs ~ week + treatment + week * treatment + (week | subject), data = BPRSL, REML = FALSE)

# print a summary of the model
summary(BPRS_ref3)
# Comment
 
```
text 

#### ANOVA for the models
```{r}

# perform an ANOVA test on the two models
 anova(BPRS_ref3, BPRS_ref2)
```
Not statistically signficant at p<0.05 


#### Draw Plots
```{r}
BPRSL$subject3<- as.numeric(BPRSL$subject)
# draw the plot of BPRSL with the observed Weight values
ggplot(BPRSL, aes(x = week, y = bprs)) +
  geom_line(aes(linetype = treatment)) +
  scale_x_continuous(name = "Time (weeks)", breaks = seq(0, 8, 1)) +
  scale_y_continuous(name = "Observed BPRS points") +
  theme(legend.position = "top")
 
```
text 


#### Title
```{r}
# Create a vector of the fitted values
Fitted <- fitted(BPRS_ref3)
# Create a new column fitted to BPRSL
BPRSL <- BPRSL %>% mutate(Fitted = Fitted)

# draw the plot of BPRSL with the Fitted values of weight
ggplot(BPRSL, aes(x = week, y = Fitted)) +
  geom_line(aes(linetype = treatment)) +
  scale_x_continuous(name = "Time ", breaks = seq(0, 8, 1)) +
  scale_y_continuous(name = "Fitted weight ") +
  theme(legend.position = "top")
 
```
text 
